Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2015-11-25 WSSSPE3/report/wg_appendix_training.tex
Appendix I:  Minor changes.

\subsection{Group Members}
\begin{itemize}
\item Nick Jones -- New Zealand eScience Infrastructure
\item Iain Larmour -- Engineering \& Physical Sciences Research Council (UK)
\item Erin Robinson -- Foundation for Earth Science
\end{itemize}

@@

While little training focuses specifically on this outcome \katznote{sustainable
software?}, a variety of training activities increase researcher awareness of
and engagement with software professionals and software engineering practices.
Research Software Engineers are being recognized as critical contributors to
high quality research; the pathway to acquire and master the relevant skills
isn't yet clear; equally those skills required by researchers in general are
also not commonly understood nor routinely developed.

The group's discussion explored a rapidly growing array of training that is seen
to contribute to sustainable software. The offerings are diverse, including:
self paced online modules focused around specific tools; single and multiple day
training workshops that raise awareness of a tool chain to support collaborative
and shared software development within a research workflow; block courses
specializing on particular methods, technologies, and applications; academic
programs at undergraduate and masters levels; doctoral training programs that in
part contain requisite skills training activities.

While some of this training focuses on applying software engineering practices
within the context of research, meeting the values and goals of research are
less often incorporated as explicit learning outcomes. With software (and
similarly, data) often the only tangible artifact of a research method or
protocol, the dependency between software applications, and the quality of
research adds complexity to the learners journey. In recognition of the longer
term investment required by researchers to integrate such skills into their
research practices, many activities are focusing on emotionally engaging
researchers and cohorts, to build a sense of shared purpose beyond the obvious
goal of technical skill acquisition.

In reviewing current training activities, the group identified a variety of
perspectives seen as useful in positioning activities in ways to better
communicate why and when best to apply each activity. Training can be
categorized on a variety of spectra, with content and delivery focused from:
programming to research; basic to advanced; technical to emotional; informal to
formal; self paced to participative. A few attempts have been made to situate a
cross section of training activities within such dimensions, creating easier
means of communicating the value of any specific opportunity and the pathways
across opportunities over time.

Evaluation of training delivery and outcomes is seen as a weakness common to
most non academic training activities. Opportunities for measuring success in
delivering training start simply with collecting a Net Promoter Score, which
lets those delivering training know whether attendees are likely to recommend
the training to others. In looking at the longer term outcomes for the learner,
frameworks such as Blooms Taxonomy and Kirkpatrick Evaluations offer possible
approaches.

@@

Coordination of these training projects will depend on buy in from a broad range
of training program and activity leaders, suggesting a key opportunity lies in
identifying and bringing together these people on a regular basis.

@@

In aiming at these opportunities it will be necessary to find the means to
support those involved in leading training activities to allocate time to
coordination activities, which will often sit beyond their current scope of
responsibility.

@@

	\item Identify programs with existing activity aimed at integrating across
	training projects.

@@

	\item Identify funding opportunities to bring together training program and
	project leaders to identify shared goals for future coordination of activities.

	\item Agree a communications plan to qualify whether programs, projects, and
	funders are interested in engaging and committing to ongoing activities.

@@

If the group moves from early stage formation into working towards shared goals,
expertise will likely be required in pedagogy and training evaluation

@@

\item 1st quarter 2016 - if willing parties identified, draft workshop proposal
and identify relevant forum, including future WSSSPE events

>>>>

\subsection{Group Members}
\begin{itemize}
\item Nick Jones -- New Zealand eScience Infrastructure
\item Iain Larmour -- Engineering \& Physical Sciences Research Council, UK
\item Erin Robinson -- Foundation for Earth Science
\end{itemize}

@@

While little training focuses specifically on this outcome \katznote{sustainable
software?}, a variety of training activities increase researcher awareness of
and engagement with software professionals and software engineering practices.
Research Software Engineers are being recognized as critical contributors to
high quality research; the pathway to acquire and master the relevant skills
is not yet clear; equally those skills required by researchers in general are
also not commonly understood nor routinely developed.

The group's discussion explored a rapidly growing array of training that is seen
to contribute to sustainable software. The offerings are diverse, including:
self-paced online modules that focused around specific tools; single and multiple day
training workshops that raise awareness of a tool chain to support collaborative
and shared software development within a research workflow; block courses
specializing on particular methods, technologies, and applications; academic
programs at undergraduate and masters levels; doctoral training programs that in
part contain requisite skills training activities.

While some of this training focuses on applying software engineering practices
within the context of research, meeting the values and goals of research are
less often incorporated as explicit learning outcomes. With software (and
similarly, data), often the only tangible artifact of a research method or
protocol, the dependency between software applications, and the quality of
research add complexity to the learners journey. In recognition of the longer
term investment required by researchers to integrate such skills into their
research practices, many activities are focusing on emotionally engaging
researchers and cohorts, to build a sense of shared purpose beyond the obvious
goal of technical skill acquisition.

In reviewing current training activities, the group identified a variety of
perspectives seen as useful in positioning activities in ways to better
communicate why and when best to apply each activity. Training can be
categorized on a variety of spectra, with content and delivery focused on:
programming to research; basic to advanced; technical to emotional; informal to
formal; self-paced to participative. A few attempts have been made to situate a
cross section of training activities within such dimensions, creating easier
means of communicating the value of any specific opportunity and the pathways
across opportunities over time.

Evaluation of training delivery and outcomes is seen as a weakness common to
most non-academic training activities. Opportunities for measuring success in
delivering training start simply with collecting a Net Promoter Score, which
let those delivering training know whether attendees are likely to recommend
the training to others. In looking at the longer term outcomes for the learner,
frameworks such as Bloom's taxonomy and Kirkpatrick's evaluation model offer possible
approaches.

@@

Coordination of these training projects will depend on buy-in from a broad range
of training program and activity leaders, suggesting a key opportunity lies in
identifying and bringing together these people on a regular basis.

@@

In aiming at these opportunities, it will be necessary to find the means to
support those involved in leading training activities to allocate time to
coordination activities, which will often sit beyond their current scope of
responsibility.

@@

	\item Identify programs with existing activities aimed at integrating across
	training projects.

@@

	\item Identify funding opportunities to bring together training programs and
	project leaders to identify shared goals for future coordination of activities.

	\item Agree on a communications plan to qualify whether programs, projects, and
	funders are interested in engaging and committing to ongoing activities.

@@

If the group moves from early-stage formation into working towards shared goals,
expertise will likely be required in pedagogy and training evaluation.

@@

\item First quarter 2016 -- if willing parties identified, draft workshop proposal
and identify a relevant forum, including future WSSSPE events.
